To derive the dispersion relation of electromagnetic wave propagating through unmagnetized plasmas, the hydrodynamic approach is exploited. Since one assume plasma response to a high frequency field, the ions are treated as a stationary, neutralizing background. Thermal motion of particles is also ignored, thus the pressure term in \ref{2.2.3} can be neglected. According to the previous assumptions, one shall solve the following set of equations,
\begin{equation}
\label{2.3.1.1}
\diffp{n_e}{t} + \nabla \cdot \left(n_e \vec{u}_e \right) = 0,
\end{equation}
\begin{equation}
\label{2.3.1.2}
m_e n_e \left[ \diffp{\vec{u}_e}{t} + \left(\vec{u}_e \cdot \nabla \right) \vec{u}_e \right] = - e n_e \vec{E}.
\end{equation}
One also needs the wave equation for the electric field \ref{1.34}. However, it is now necessary to include the current density due to motion of charged particles in plasma. The hydrodynamic equations are coupled with the Maxwell's equations via $ \vec{J} = q_s n_s \vec{u}_s $, thus the wave equation yields the following form,
\begin{equation}
\label{2.3.1.3}
\laplace{\vec{E}} - \frac{1}{c^{2}} \diffp[2]{\vec{E}}{t} = - \mu_0 e n_e \diffp{\vec{u}_e}{t},
\end{equation}

The system of equations above will be linearized using the methods of perturbation theory. Consider a small perturbations from the stationary state,
\begin{equation}
\label{2.3.1.4}
n_{e} = n_{e0} + \delta n_{e}, \quad \vec{u}_{e} = \vec{u}_{e0} + \delta \vec{u}_{e}, \quad \vec{E} = \vec{E}_{0} + \delta \vec{E},
\end{equation}
where $ \vec{u}_{e0} $ and $ \vec{E}_{0} $ are obviously identically equal to zero vector. After substituting perturbed quantities \ref{2.3.1.4} into initial system of equations and performing Fourier transform one obtains
\begin{equation}
\label{2.3.1.5}
\delta n_{e} = \mathrm{i} \frac{n_{e0}}{\omega} \vec{k} \cdot \delta \vec{u}_{e},
\end{equation}
\begin{equation}
\label{2.3.1.6}
\delta \vec{u}_e = -\i \frac{e}{\omega m_e} \delta \vec{E},
\end{equation}
\begin{equation}
\label{2.3.1.7}
\delta \vec{E} = \i \frac{\omega \, e \, n_{e0}}{\varepsilon_0 \left(\omega^2 - c^2 k^2 \right)} \delta \vec{u}_e.
\end{equation}
Note that the equation for density perturbation \ref{2.3.1.5} may be ignored. Eliminating $ \delta \vec{u}_{e} $ from the equation \ref{2.3.1.7} one gets the equation for perturbation of the electric field,
\begin{equation}
\label{2.3.1.8}
\left(\omega^2 - \omega_{pe}^2 - c^2 k^2 \right) \delta \vec{E} = 0. 
\end{equation}

The dispersion relation can be immediately obtained from the equation \ref{2.3.1.8}. To describe how the electromagnetic waves propagates through given medium, it is useful to introduce the index of refraction $ N \left( \omega \right) =  c k \left( \omega \right) / \omega $. Consequently, the dispersion relation may be rewritten as
\begin{equation}
\label{2.3.1.9}
N^{2} = 1 - \left(\frac{\omega_{pe}}{\omega}\right)^2.
\end{equation} 
The important properties of the waves are distinguished by their cut-offs ($ N \rightarrow 0 $) and resonances ($ N \rightarrow \infty $). In the vicinity of the resonance there is a total absorption, at a cut-off frequency there is a total reflection of incident waves.

In the case of the electromagnetic wave propagating through unmagnetized plasma, it might be clearly seen that there are no resonances. On the other hand, the equation \ref{2.3.1.9} exhibits cut-off and the corresponding frequency (including ions) is given by the following expression,
\begin{equation}
\label{2.3.1.10}
\omega = \sqrt{\omega_{pe}^{2} + \omega_{pi}^{2}}
\end{equation}
The condition \ref{2.3.1.10} occurs at the so-called critical plasma density $ n_c $. Note that the electromagnetic wave passing through plasma with densities larger than $ n_c $ is exponentially damped.